---
title: "Task 4- Naples Reef Purple Sea Urchins"
author: "Deanna Mireles"
date: "10/28/2021"
output: 
  html_document: 
    toc: yes
    theme: yeti
---

## Overview

This code uses data for red and purple sea urchins and in this task X. The data was collected in Santa Barbara by Daniel Reed in 2020 as part of LTER.

**Data citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Setup

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
```

Here I read in the sea urchin data using the here() function.

```{r}
sea_urchin <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

For reference, the first 6 lines of the dataset are shown below:

```{r}
head(sea_urchin)
```

## Data Wrangling

```{r}

```

## Data Visualization- Exploratory histogram and qq-plot

```{r}

```


## Data Visualization- Summary table

```{r}

```



## End Task